### 2019-09-20 Manuscript/Supplementary_Information/Supplementary_Information.tex

The silicate weathering component of the GEOCLIM model as been improved compare to the previous published version (Goddéris et al, 2017 natgeo). The new module implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile. We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock. In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, oras a physically weathered particle once it reaches the top. We use the DynSoil implementation of the \cite{Gabet2009a} model which integrates a climatic dependence on the chemical weathering that occurs within the regolith using the formulation of \cite{West2012a}. The transient time-varying version of this regolith model is described by the following system of three equations which form the core of the new DynSoil module of GEOCLIM:

@@

\noindent
$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}. This formulation is close to the large-scale BQART model of \citet{Syvitski2007a}, but without a temperature dependence and the replacement of elevation with slope, resulting in functional form that more closely resembles the stream power law \citep{Davy2000a}. This formulation and these exponent values are supported by various compilations (e.g. \citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is a variability in the proportionality constant that is difficult to capture at a global scale.

@@

\noindent
$x(z)$ is the abundance profile of primary phases inside the regolith, varying with height upward from the base of the regolith as shown in Figure \ref{fig:regolith_schematic}. Setting $z$ equal to the regolith thickness ($h$) gives $x_s$ which is the proportion of primary phases remaining at the top of the regolith column. This value controls the overall integrated chemical weathering rate out of the regolith.

@@

We restricted the calculation of the weathering fluxes to the flux of dissolved Mg and Ca originating from continental silicate weathering, as they are the sole contributors to the long term consumption of atmospheric CO2. To calculate this flux associated with chemical weathering, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig. \ref{fig:regolith_schematic}). Previous implementations of GEOCLIM have used bulk continental crust values across all exposed land, using a concept of diffuse lithology (it was assume that the weathering rates of shield rocks, basaltic rocks, and carbonate rocks was the same for each continental grid element, when this grid element was submitted to the same climatic conditions). Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, the diffuse lithology cannot be used anymore, we have implemented a realistic lithologically-resolved version of the model.

@@

The majority of the original 78,000 parameter combinations produce \COtwo consumption maps with poor fits to the measured watershed data (Fig. \ref{fig:W_vs_r2}). Given that the coefficient of determination ($r^{2}$) is calculated using equation \ref{eq:10} rather than fitting a linear model, many of the combinations associated with particularly poor fits result in negative $r^{2}$ values (Fig. \ref{fig:W_vs_r2}). However, given the right permutation of parameters, GEOCLIM produces \COtwo consumption maps that fit the measured watershed data reasonably well. We therefore eliminate all parameter combinations that produce a $r^{2}\textless0.4$, which leaves 3,339 unique parameter combinations (Figs. \ref{fig:W_vs_r2} and \ref{fig:r2_cross_plot}).

The mean global CO$_{2}$ consumption calculated from these 3,339 parameter combinations is 5.4$\times10^{12}$~mol/yr, with a standard deviation of 0.9$\times10^{12}$~mol/yr. These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig. \ref{fig:W_vs_r2}B).

@@

We use temperature and runoff fields from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; available for download at \url{https://nomads.gfdl.noaa.gov/dods-data/gfdl_cm2_0/}) for the climate model component of GEOCLIM. These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca. 1860 at a 2.0\degrees $\times$ 2.5\degrees resolution. In the ``1860 control'' experiment, forcing agents representative of conditions ca. 1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium. \pCOtwo in 1860 is assumed to be 286~ppm. In the ``+1\%/yr to 2$\times$'' experiment, initial conditions are taken from the ``1860 control'' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value. \pCOtwo is then held constant until the end of the 280 year experiment. All non-\COtwo forcing agents are held constant. The ``+1\%/yr to 4$\times$'' experiment is identical to the ``+1\%/yr to 2$\times$'' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value. \pCOtwo is then held constant for 160 years. We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/ING_fracs.jpg}
    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of Indonesia and New Guinea (I\&NG) used to force GEOCLIM. Each column represents a tested scenario. Each row represents a lithologic category. Solid black lines show present-day shorelines.}
    \label{fig:ING_fracs}
\end{figure}

We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y (described in detail in \textit{Paleoshorelines}). To generate the lithologic map used in the ``removed I\&NG'' scenario (Fig. \ref{fig:ING_fracs}), we remove all land associated with arc-continent collision in I\&NG. To generate the lithologic map used in the ``reduced I\&NG'' scenario (Fig. \ref{fig:ING_fracs}), we take the estimated paleoshorelines of I\&NG for 15~Ma, then remove all land that falls outside of these bounds. For this scenario, we do not add any land (i.e. the Sunda Shelf) that is not exposed today to the lithologic map, and the slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of I\&NG for 15~Ma.

The lithologic map used in the ``reduced I\&NG + Sunda Shelf'' scenario (Fig. \ref{fig:ING_fracs}) is identical to that used in the ``reduced I\&NG'' scenario, with the addition of the Sunda Shelf. However, the assignment of lithology, runoff, and slope for pixels on the Sunda Shelf is not trivial, given that the shelf is not currently exposed. As discussed in \textit{Paleoshorelines}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust. Basement highs between Malaysia/Sumatra and Borneo are typically granitic and bounded by basins dominated by shallow marine clastic fill. However, basement highs are less common in the Gulf of Thailand. We therefore assign all pixels of the Sunda Shelf to be 50\% felsic and 50\% sediment, except in the Gulf of Thailand where pixels are assigned to be 100\% sediment. For the slope, we take the mean slope value (0.0038~m/m) of relatively flat land on eastern Sumatra, and assign this value to all pixels of the Sunda Shelf.  For the runoff, we linearly interpolate along latitude bands between present-day land pixels to obtain runoff values for the presently-submerged Sunda Shelf. This approach preserves large N-S variation in runoff associated with the large-scale Hadley circulation and therefore should be reasonable to first-order, although it does not accurately capture the smaller-scale interactions between land and ocean/atmosphere circulation.

@@

Another limitation associated with using a modern-day climate model, is that in the scenarios with varying shorelines, we hold the slope and lithology to be the same as the present-day in each pixel. This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of any emerged pixel the same. Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order. However, implementing high-resolution climate models with both varying shorelines and change in paleogeographic position for these scenarios is valuable future direction, although it would require a better knowledge of the past continental topography, which is always challenging. Additionally, such modeling efforts could give insight into whether associated changes in ocean and atmosphere circulation would have significantly modulated precipitation in I\&NG.  Given that precipitation in this region is significantly driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for I\&NG. However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.

@@

\section*{Paleoshoreline reconstruction and geological synthesis}

@@

Paleogeographic reconstructions of Peninsular Malaysia suggest that it has been largely exposed over the last 20~Ma \citep{Hall2013b, Hall2002a}. Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were also emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}. Similarities between the terrestrial biotas of Borneo and mainland Southeast Asia confirm the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}. Eocene through Miocene rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf \citep{Morley2013a}. Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially enclosed inland seas and extensive brackish or saline wetland environments. Palynological analysis suggests widespread swamp environments persisting through the Late Miocene \citep{Morley2013a}. Basement highs (e.g. the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill \citep{Darmadi2007a}. Most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20~Ma onwards \citep{Hall2013b, Madon2013a}, although it is omitted in the 5 Ma paleogeographic model of \citet{Molnar2015a}.

@@

Separated from the Paleozoic continental core of SW Borneo by the Lupar Line suture zone, the northern portion of Borneo (Sabah) is underlain by ophiolitic basement that extends to Palawan \citep{Hall2008a, Ilao2018a}. The Sarawak Basin and NW Borneo trough offshore Sabah host $\textgreater$10~km-thick Neogene sedimentary sequences, indicative of the extent and duration of exhumation in northern Borneo \citep{Hall2008a}. The Late Oligocene/Early Miocene Sabah Orogeny resulted in the uplift of both Sabah and southern Palawan \citep{Hall2013a}, as well as the obduction of the Palawan ophiolite and Telupid ophiolite of northern Sabah. Under-thrusting of thinned passive-margin continental crust beneath these suprasubduction ophiolites resulted in Early to Middle Miocene exhumation and offshore unconformities \citep{Hall2008a}. Early Miocene sediments in northern Sabah have a provenance from Palawan \citep{Suggate2014a}, indicating that Palawan, like northern Borneo, experienced uplift during the Sabah orogeny. However, initiation of back-arc extension in the Sulu Sea around 19~Ma \citep{Hall2013a} resulted in the subsidence of the eastern Palawan Mountains, eliminating or reducing the role of Palawan as a sediment source for the Borneo trough. The 14~Ma Capoas Granite on Palawan and the 7.5~Ma Kinabalu Granite in Sabah are interpretted to be the result of crustal thinning related to pulses of regional backarc extension \citep{Hall2013a}. Thermochronological data from the Sabah highlands suggest extremely rapid uplift (7~km/m.y.) and exhumation during the latest Miocene and early Pliocene, resulting in the recent formation of Mt. Kinabalu, the tallest peak in Borneo \citep{Cottam2013a}.

@@

For the paleoshorelines of Sulawesi, we largely followed the recent stratigraphic compilation and paleoshorelines compilation of \citet{Nugraha2018a}. Geographically, Sulawesi can be divided into a central highland region flanked by North, South, Southeast, and East Arms. These arms have high relief ($\textgreater$3~km) that are separated by deep basins and broadly define Sulawesi's seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \citep{Hamilton1979a, Katili1978a}. Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system \citep{Parkinson1998b, Hennig2016a, Hall2017a, Hennig2017a}. After collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed \citep{Polve1997a, Elburg1998a} that was connected to the paleo-Sunda Arc \citep{Hall2002b}, also referred to as the Great Indonesian arc \citep{Harris2006a}.  The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting (\citealp{Monnier1995a}, but see \citealp{Kadarusman2004a} for an alternative view), and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32--28~Ma; \citealp{Parkinson1998a}) and thrust to the east above a west dipping slab (East Sulawesi block of \citealp{Villeneuve2001a}). During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing ophiolite emplacement in West and SE Sulawesi, including Buton \citep{Smith1991a, Bergman1996a}.

@@

Along the Papuan Peninsula of Eastern New Guinea, the PUB was obducted above a north dipping slab during Oligocene arc-continent collision between Australian continental fragments and the Melanesian arc, but remained largely subaqueous until the Miocene \citep{Davies1971a, Davies1984a, vanUfford2005a}. Miocene to present arc-continent collision in New Guinea has progressed from west to east. Exhumation of the Central Range accelerated over the past 4~m.y. which is interpretted to be the result of slab-breakoff and buoyant uplift \citep{Cloos2005a}. A change to left-lateral lateral motion on the northern coast of New Guinea at this time, caused the exhumation of ophiolites both along the coast \citep{Monnier1999a} and on the islands Obi \citep{Ali2001a} and Halmahera \citep{Hall1988a, Ballantyne1992a}. In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~Ma \citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.

@@

Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate \citep{Harris2006a}. From ca. 20--10~Ma, the majority of Sumatra and nearly all of Java were submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf \citep{Hall2009a, Hall2013b}. Six major islands were subaerially exposed before 5~Ma: Sumatra, Belitung, Bangka, Java, Bali, and Flores. However, most of Sumatra and Java were elevated above sea level and emerged to their present exposures only since 5~Ma \citep{Hall2009a, Hall2013b}. Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{AudleyCharles2004a, Harris2006a, Hall2013b}. In Timor and Sumba, arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations $\textgreater$1~km above sea level with estimated average uplift rates of 1.5~km/m.y. \citep{AudleyCharles1986a}.

>>>>

The silicate weathering component of the GEOCLIM model has been improved compared to the previously published version \citep{Godderis2017b}. The new component implements the model of \cite{Gabet2009a} for the development of a chemically weathered profile. We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock. In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, or as a physically weathered particle once it reaches the top. We use the DynSoil implementation of the \cite{Gabet2009a} model which integrates a climatic dependence on the chemical weathering that occurs within the regolith using the formulation of \cite{West2012a}. The transient time-varying version of this regolith model is described by the following system of three equations which form the core of the new DynSoil component of GEOCLIM:

@@

\noindent
$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 as in \citet{Maffre2018a}. This formulation is close to the large-scale BQART model of \citet{Syvitski2007a}, but without a temperature dependence and the replacement of elevation with slope, resulting in functional form that more closely resembles the stream power law \citep{Davy2000a}. This formulation and these exponent values are supported by various compilations (e.g. \citealp{Lague2013a}) although \citet{Lague2013a} also suggests that there is variability in the proportionality constant that is difficult to capture at a global scale.

@@

\noindent
$x(z)$ is the abundance profile of primary phases inside the regolith, varying with height upward from the base of the regolith as shown in Figure \ref{fig:regolith_schematic}. Setting $z$ equal to the regolith thickness ($h$) gives $x_s$ which is the proportion of primary phases remaining at the top of the regolith column.

@@

We restricted the calculation of the weathering fluxes to the flux of dissolved Ca+Mg originating from continental silicate weathering, as they are the sole contributors to the long-term consumption of atmospheric \COtwo. To calculate this flux, we need to assign the concentration of Ca and Mg ($\chi_{CaMg}$) within the unweathered bedrock (Fig. \ref{fig:regolith_schematic}). Previous implementations of GEOCLIM have used bulk continental crust values across all exposed land, using the concept of ``diffuse lithology'' (it was assumed that the weathering rates of all rocks was the same for each continental grid element, provided that the grid element was submitted to the same climatic conditions). Given that the hypothesis we seek to test involves the varying concentration of cations in different lithologies, the diffuse lithology cannot be used anymore, and instead we have implemented a more realistic lithologically-resolved version of the model.

@@

The majority of the original 78,000 parameter combinations produce \COtwo consumption maps with poor fits to the measured watershed data (Fig. \ref{fig:W_vs_r2}). Given that the coefficient of determination ($r^{2}$) is calculated using equation \ref{eq:10} rather than fitting a linear model, many of the combinations associated with particularly poor fits result in negative $r^{2}$ values (Fig. \ref{fig:W_vs_r2}). However, given the right permutation of parameters, GEOCLIM produces \COtwo consumption maps that fit the measured watershed data reasonably well. We therefore eliminate all parameter combinations that produce a $r^{2}<0.4$, which leaves 3,339 unique parameter combinations (Figs. \ref{fig:W_vs_r2} and \ref{fig:r2_cross_plot}).

Global CO$_{2}$ consumption calculated from these 3,339 parameter combinations ranges from 2.7$\times10^{12}$~mol/yr to 7.1$\times10^{12}$~mol/yr, with a mean of 5.4$\times10^{12}$~mol/yr. These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig. \ref{fig:W_vs_r2}B).

@@

We use temperature and runoff fields from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; available for download at \url{https://nomads.gfdl.noaa.gov/dods-data/gfdl_cm2_0/}) for the climate model component of GEOCLIM. These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca. 1860 at a 2.0\degrees $\times$ 2.5\degrees resolution. In the ``1860 control'' experiment, forcing agents representative of conditions ca. 1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium. \pCOtwo in 1860 is assumed to be 286~ppm. In the ``+1\%/yr to 2$\times$'' experiment, initial conditions are taken from the ``1860 control'' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value. \pCOtwo is then held constant until the end of the 280 year experiment. All non-\COtwo forcing agents are held constant. The ``+1\%/yr to 4$\times$'' experiment is identical to the ``+1\%/yr to 2$\times$'' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value. \pCOtwo is then held constant for 160 years. We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively. Temperature and runoff fields for any arbitrary \pCOtwo value between 286 and 1144~ppm is obtained by interpolating between the three modeled \pCOtwo levels.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/ING_fracs.jpg}
    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of Indonesia and New Guinea (I\&NG) used to force GEOCLIM. Only the 15~Ma scenarios are shown here, but the 10 and 5~Ma scenarios would be similar, using the 10 and 5~Ma shorelines shown in Figure 1 of the main text instead. Each column represents a tested scenario. Each row represents a lithologic category. Solid black lines show present-day shorelines.}
    \label{fig:ING_fracs}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Figures/scenario_pCO2_all.pdf}
    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text. For each of the ten scenarios, each point represents an estimate from one of the 3,339 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched \COtwo consumption estimated from watershed data (see SI). The box encloses the middle 50\% of the \pCOtwo estimates (i.e. the interquartile range), and the notch represents the median with its 95\% confidence interval. The whiskers extend to 1.5$\times$ the interquartile range, or to the minimum/maximum of the estimates if no estimates are below/above this range.}
    \label{fig:scenario_pCO2_all}
\end{figure}

We use geological data to quantify the changes in the area of I\&NG over the past 20~m.y (described in detail in \textit{Paleoshoreline Reconstruction and Geological Synthesis}). To generate the lithologic map used in the ``removed I\&NG'' scenario (Fig. \ref{fig:ING_fracs}), we remove all land associated with arc-continent collision in I\&NG. To generate the lithologic map used in the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig. \ref{fig:ING_fracs}), we take the estimated paleoshorelines of I\&NG for 15, 10, and 5~Ma, then remove all land that falls outside of these bounds. The slope and lithologic classification of each pixel is identical to that of the present day, provided that these pixels fall within the estimated paleoshorelines of I\&NG for their respective time slice. We then add land that is not exposed today (i.e. the Sunda Shelf). However, the assignment of lithology, runoff, and slope for pixels on the Sunda Shelf is not trivial, given that the shelf is not currently exposed. As discussed in \textit{Paleoshoreline Reconstruction and Geological Synthesis}, the Sunda Shelf is a flat-lying, relatively stable platform of continental crust. Basement highs between Malaysia/Sumatra and Borneo are typically granitic and bounded by basins dominated by shallow marine clastic fill. However, basement highs are less common in the Gulf of Thailand. We therefore assign all pixels of the Sunda Shelf to be 50\% felsic and 50\% sediment, except in the Gulf of Thailand where pixels are assigned to be 100\% sediment. For the slope, we take the mean slope value (0.0038~m/m) of relatively flat land on eastern Sumatra, and assign this value to all pixels of the Sunda Shelf.  For the runoff, we linearly interpolate along latitude bands between present-day land pixels to obtain runoff values for the presently-submerged Sunda Shelf. This approach preserves large N-S variation in runoff associated with the large-scale Hadley circulation and therefore should be reasonable to first-order, although it does not accurately capture the smaller-scale interactions between land and ocean/atmosphere circulation.

To explore the sensitivity of our results to the inclusion of the Sunda Shelf, we also tested ``reduced I\&NG'' scenarios for 15, 10, and 5~Ma, which are identical to that of the ``reduced I\&NG + Sunda Shelf'' scenarios without the addition of the Sunda Shelf. We find that the estimated steady state \pCOtwo's for the ``reduced I\&NG'' scenarios are only marginally higher than those of the ``reduced I\&NG + Sunda Shelf'' scenarios (Fig. \ref{fig:scenario_pCO2_all}), suggesting that our results are largely insensitive to the inclusion of the Sunda Shelf. This insensitivity can be attributed to the low Ca+Mg concentrations of felsic and sediment lithologies in conjunction with the low slope values distributed across the shelf.

@@

Another limitation associated with using a modern-day climate model, is that in the scenarios with varying shorelines, we hold the slope and lithology to be the same as the present-day in each pixel. This means that while we are changing the amount of emergent land, we are not changing the latitudinal or longitudinal position and are keeping the slope of any emerged pixel the same. Given that similar distributions of lithology and slope are reasonable in the past and that latitudinal translation has been relatively minor, we consider this simplification to be reasonable to first-order. However, implementing high-resolution climate models with both varying shorelines and changes in paleogeographic position for these scenarios is a valuable future direction. Additionally, such modeling efforts could give insight into whether associated changes in ocean and atmosphere circulation would have significantly modulated precipitation in I\&NG.  Given that precipitation in this region is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water, the current approach is a reasonable approximation for I\&NG. However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable.

@@

\section*{Paleoshoreline Reconstruction and Geological Synthesis}

@@

Paleogeographic reconstructions of Peninsular Malaysia suggest that it has been largely exposed over the last 20~Ma \citep{Hall2013b, Hall2002a}. Although the majority of the Sunda Shelf is currently submerged, large portions of this flat-lying, relatively stable platform of continental crust were also emergent throughout the Neogene, and were repeatedly subaerially exposed during the Pleistocene as recently as the Last Glacial Maximum \citep{Hall2002a}. Similarities between the terrestrial biotas of Borneo and mainland Southeast Asia confirm the existence of land bridges on the Sunda Shelf throughout the Miocene \citep{Moss1998a}. Eocene through Miocene rifting of the South China Sea resulted in subsidence throughout the Sunda Shelf \citep{Morley2013a}. Grabens associated with this subsidence became freshwater rift lakes that later transitioned to partially enclosed inland seas and extensive brackish or saline wetland environments. Palynological analysis suggests widespread swamp environments persisting through the Late Miocene \citep{Morley2013a}. Basement highs (e.g. the Natuna Arch, currently subaerially exposed as the largely granitic island of Natuna off SW Borneo, as well as the granitic Tin Islands south of the Malay Peninsula) are typically bounded by Paleogene to Neogene basins dominated by shallow marine clastic fill \citep{Darmadi2007a}. Most paleogeographic reconstructions of the region incorporate some degree of exposure of the Sunda Shelf from 20~Ma onwards \citep{Hall2013b, Madon2013a}, although it is omitted in the 5~Ma paleogeographic model of \citet{Molnar2015a}.

@@

Separated from the Paleozoic continental core of SW Borneo by the Lupar Line suture zone, the northern portion of Borneo (Sabah) is underlain by ophiolitic basement that extends to Palawan \citep{Hall2008a, Ilao2018a}. The Sarawak Basin and NW Borneo trough offshore Sabah host $>$10~km-thick Neogene sedimentary sequences, indicative of the extent and duration of exhumation in northern Borneo \citep{Hall2008a}. The Late Oligocene/Early Miocene Sabah Orogeny resulted in the uplift of both Sabah and southern Palawan \citep{Hall2013a}, as well as the obduction of the Palawan ophiolite and Telupid ophiolite of northern Sabah. Under-thrusting of thinned passive-margin continental crust beneath these suprasubduction ophiolites resulted in Early to Middle Miocene exhumation and offshore unconformities \citep{Hall2008a}. Early Miocene sediments in northern Sabah have a provenance from Palawan \citep{Suggate2014a}, indicating that Palawan, like northern Borneo, experienced uplift during the Sabah orogeny. However, initiation of back-arc extension in the Sulu Sea around 19~Ma \citep{Hall2013a} resulted in the subsidence of the eastern Palawan Mountains, eliminating or reducing the role of Palawan as a sediment source for the Borneo trough. The 14~Ma Capoas Granite on Palawan and the 7.5~Ma Kinabalu Granite in Sabah are interpreted to be the result of crustal thinning related to pulses of regional backarc extension \citep{Hall2013a}. Thermochronological data from the Sabah highlands suggest extremely rapid uplift (7~km/m.y.) and exhumation during the latest Miocene and early Pliocene, resulting in the recent formation of Mt. Kinabalu, the tallest peak in Borneo \citep{Cottam2013a}.

@@

For the paleoshorelines of Sulawesi, we largely followed the recent stratigraphic compilation and paleoshorelines compilation of \citet{Nugraha2018a}. Geographically, Sulawesi can be divided into a central highland region flanked by North, South, Southeast, and East Arms. These arms have high relief ($>$3~km) that are separated by deep basins and broadly define Sulawesi's seven tectonic provinces: 1) the West Sulawesi magmatic arc of the South Arm, 2) the Central Sulawesi metamorphic Belt, 3) the Sangihe arc of the North Arm, 4) the East Sulawesi Ophiolite of the East Arm, 5) the southeast metamorphic belt of the Southeast Arm, and the microcontinental blocks of 6) Banggai-Sula and 7) Buton-Tukang Besi \citep{Hamilton1979a, Katili1978a}. Like SW Borneo, the North and South Arms and much of Central Sulawesi are underlain by continental blocks that rifted off of the Australian-Birds Head margin in the Jurassic and collided in the Cretaceous with Eurasian basement of Sundaland as part of the Woyla Arc system \citep{Parkinson1998b, Hennig2016a, Hall2017a, Hennig2017a}. After collision, subduction polarity reversed and a Cretaceous to Miocene SE-facing volcanic arc developed \citep{Polve1997a, Elburg1998a} that was connected to the paleo-Sunda Arc \citep{Hall2002b}, also referred to as the Great Indonesian arc \citep{Harris2006a}.  The ophiolites of Sulawesi were generated in this arc system in a back arc to intra-arc setting (\citealp{Monnier1995a}, but see \citealp{Kadarusman2004a} for an alternative view), and the largest fragments of the East Sulawesi ophiolite were detached from their metamorphic sole in the late Oligocene (32--28~Ma; \citealp{Parkinson1998a}) and thrust to the east above a west dipping slab (East Sulawesi block of \citealp{Villeneuve2001a}). During the Early to Middle Miocene, the Tukang Besi-Buton block began to collide obliquely with the southeastern end of the Sunda Arc, causing ophiolite emplacement in West and SE Sulawesi, including Buton \citep{Smith1991a, Bergman1996a}.

@@

Along the Papuan Peninsula of Eastern New Guinea, the PUB was obducted above a north dipping slab during Oligocene arc-continent collision between Australian continental fragments and the Melanesian arc, but remained largely subaqueous until the Miocene \citep{Davies1971a, Davies1984a, vanUfford2005a}. Miocene to present arc-continent collision in New Guinea has progressed from west to east. Exhumation of the Central Range accelerated over the past 4~m.y. which is interpreted to be the result of slab-breakoff and buoyant uplift \citep{Cloos2005a}. A change to left-lateral lateral motion on the northern coast of New Guinea at this time, caused the exhumation of ophiolites both along the coast \citep{Monnier1999a} and on the islands Obi \citep{Ali2001a} and Halmahera \citep{Hall1988a, Ballantyne1992a}. In eastern New Guinea, progressive jamming of the north-dipping subduction zone has caused major uplift over the past 4~Ma \citep{vanUfford2005a}, which is well dated with a change in provenance from continental to volcanic detritus \citep{Abbott1994a} and thermochronology \citep{Hill1989a}.

@@

Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate \citep{Harris2006a}. From ca. 20--10~Ma, the majority of Sumatra and nearly all of Java were submerged, although Bangka and Belitung were exposed in addition to a portion of the greater Sunda Shelf \citep{Hall2009a, Hall2013b}. Six major islands were subaerially exposed before 5~Ma: Sumatra, Belitung, Bangka, Java, Bali, and Flores. However, most of Sumatra and Java were elevated above sea level and emerged to their present exposures only since 5~Ma \citep{Hall2009a, Hall2013b}. Most of the non-volcanic islands of the Outer Banda Arc emerged after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \citep{AudleyCharles2004a, Harris2006a, Hall2013b}. In Timor and Sumba, arc-continent collision resulted in rapid uplift of deep marine sedimentary rocks to elevations $>$1~km above sea level with estimated average uplift rates of 1.5~km/m.y. \citep{AudleyCharles1986a}.
